---
title: "DSM and Heterogeneity"
author: "Jihyeon bae"
date: "2024-01-03"
output: html_document
---


# Introduction

## Background and Goals

- Authoritarian Intergovernmental Organizations (AIGO): formal IGOs that largely comprise authoritarian states 
  - Creating aggregate level variable based on country-level data
  - Creating indices for robustness check

- Finding distinctive patterns of AIO compared to democratic counterparts

# Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("pacman")
p_load(plyr, dplyr, tidyr, ggplot2, tidyverse, RColorBrewer, readxl,
       readr, haven, countrycode,
       rqog, igoR, modelsummary, knitr, kableExtra, flextable)

```

# Loading Raw data
```{r, echo=FALSE, warning=FALSE}

#Qualities of Government
qog<-rqog::read_qog(which_data = "standard", data_type = "time-series") 
qog<-qog%>%
  filter(year>=1970 & year<2022) 

#Measuring International Authority
#MIA <- read_dta("~/Desktop/Regime-Heterogeneity-and-DSM/data/raw/DP_May 2021.dta")

#Downloaading the updated version
MIA <- read_csv("~/Desktop/Regime-Heterogeneity-and-DSM/data/raw/MIA_updated.csv")
#Correlates of War
COW<-igoR::igo_year_format3 %>%
  filter(year>=1970)
```

---

# Country-level variables

```{r}
country <-qog %>%
  dplyr::select(cname, year, ccodecow, 
         # World Development Indicators
          wdi_gdpcapcon2015, wdi_pop, wdi_trade, 
         # Rule of Law
          fh_rol, wbgi_rle,
      # Security variables
          wbgi_pve, # political stability
          cspf_sfi, #fragility index
          atop_number, 
      # Democracy Indices 
          vdem_polyarchy, vdem_libdem, vdem_partipdem,
      # Democracy Indices (Alternatives for robustness check)
          p_polity2, bmr_dem, ht_regtype1,
      # Globalization Index
           kofgi_dr_eg=dr_eg,
           kofgi_dr_ig=dr_ig,
           kofgi_dr_pg=dr_pg,
           kofgi_dr_sg=dr_sg,
      # colonial legacy
           ht_colonial,
      # Political Terror Scale
         pts_ptsa=gd_ptsa, 
         pts_ptsh=gd_ptsh, 
         pts_ptss=gd_ptss,
      # NELDA
         nelda_mbbe, #Media bias before election
         nelda_oa) %>%  # was opposition allowed
  dplyr::mutate(wdi_log_gdpcapcon2015=log(wdi_gdpcapcon2015),
         wdi_log_pop=log(wdi_pop),
         wdi_log_trade=log(wdi_trade))%>%
  dplyr::rename(ccode=ccodecow)%>%
  dplyr::relocate(ccode, cname, year)

country <- country %>%
  mutate(ht_colonial=ifelse(ht_colonial>1, 1, 0),
         ht_regtype1=case_when(ht_regtype1==9 ~ 5, 
                            ht_regtype1==99 ~ 6,
                            ht_regtype1==100 ~ 7,
                            TRUE ~ ht_regtype1))
  
```

```{r}
#Check rows that have NA values for VARIABLE_OF_INTEREST
country[is.na(country$ccode),]

#Fill in NAs
country$ccode[country$cname == "Ethiopia"] <- 530
country$ccode[country$cname == "Germany"] <- 255
country$ccode[country$cname == "Yemen Democratic"] <- 680
country$ccode[country$cname == "Yemen"] <- 679
country$ccode[country$cname == "Sudan"] <- 625
country$ccode[country$cname == "Cyprus"] <- 352
country$ccode[country$cname == "Pakistan"] <- 770
country$ccode[country$cname == "Vietnam, North"] <- 816
country$ccode[country$cname == "Vietnam, South"] <- 817
country$cname[country$cname == "USSR"] <- "Russian Federation (the)"


#Run below code to check year-availability for VARIABLE_OF_INTEREST
#summary(country[!is.na(country$VARIABLE_OF_INTEREST),])

country<-country%>%
  filter(!is.na(ccode))

```

---

## Transposing IGO-level data into country-level data

- Pivot longer so that row represents country
```{r}
COW <- COW %>%
  dplyr::rename(cow_igocode = ionum)%>%
  dplyr::select(-c(igocode, version, accuracyofpre1965membershipdates,sourcesandnotes, imputed)) %>%
  dplyr::relocate(cow_igocode, ioname, year, political, social, economic)%>%
  pivot_longer(c(`afghanistan`:`zimbabwe`),
                      names_to="country",
                      values_to="membership")%>%
  dplyr::filter(membership==1) #member states only

```

## Identify ambiguous country names
```{r}
COW <- COW %>%
  mutate(country = recode(country, "austriahungary" = "Austria-Hungary",
                          "domrepublic"="Dominican Republic",
                          "etimor"="East Timor",
                          "hessegrand"="Hesse Grand Ducal",
                          "micronesiafs"="Federated States of Micronesia",
                          "nokorea"="North Korea",
                          "soafrica"="South Africa",
                          "sokorea"="South Korea",
                          "stlucia"="St. Lucia",
                          "wgermany"="German Federal Republic",
                          "syemen"="Yemen People's Republic",
                          .default = country))

#Attaching Country numeric code to character values
COW$ccode<-countrycode(COW$country, 
              origin='country.name', 
              destination='cown', 
              warn = TRUE)

COW<-COW%>%
  filter(!is.na(ccode))
```


## Extracting member-state information for each IGO
```{r}
igo_master <- COW %>%
  dplyr::left_join(country, by=c("ccode", "year"))%>%
  dplyr::select(-c(orgname, longorgname, membership))

igo_master <- igo_master %>%
  group_by(cow_igocode, year)%>%
  mutate(gdp_share=wdi_gdpcapcon2015 / sum(wdi_gdpcapcon2015, na.rm=TRUE),
         poly_share=vdem_polyarchy / sum(vdem_polyarchy, na.rm=TRUE))
```



# Variable Generation
## summarising state level info to igo level
```{r}
igo <- igo_master %>%
  dplyr::group_by(cow_igocode, ioname, year) %>%
  dplyr::summarise(
    #average democracy scores
    polyarchy = mean(vdem_polyarchy, na.rm = TRUE),
    polyarchy_median = median(vdem_polyarchy, na.rm = TRUE),
    partipdem = mean(vdem_partipdem, na.rm=TRUE),
    partipdem_median = median(vdem_partipdem, na.rm=TRUE),
    libdem = mean(vdem_libdem, na.rm = TRUE),
    libdem_median = median(vdem_libdem, na.rm = TRUE),
    polity = mean(p_polity2, na.rm = TRUE),
    polity_median = median(p_polity2, na.rm = TRUE),
    regtype=mean(ht_regtype1, na.rm=TRUE),
    regtype_median=median(ht_regtype1, na.rm=TRUE),
    #economic variables
    gdp_cap = mean(wdi_log_gdpcapcon2015, na.rm = TRUE),
    population=mean(wdi_log_pop, na.rm=TRUE),
    trade=mean(wdi_log_trade, na.rm=TRUE),
    globalization=mean(kofgi_dr_ig, na.rm=TRUE),
    globalization_political=mean(kofgi_dr_pg, na.rm=TRUE),
    alliances = mean(atop_number, na.rm = TRUE),
    # Governance
    fh_rol=mean(fh_rol, na.rm = TRUE),
    wbgi_rle=mean(wbgi_rle, na.rm=TRUE),
    wbgi_pve=mean(wbgi_pve, na.rm=TRUE),
    cspf_sfi=mean(cspf_sfi, na.rm=TRUE),
    ptsa=mean(pts_ptsa, na.rm=TRUE), 
    #characteristics 
    number = n(),
    trade = mean(wdi_log_trade, na.rm = TRUE),
    percentage = sum(bmr_dem, na.rm = TRUE) / number,
    political = mean(political, na.rm = TRUE),
    social = mean(social, na.rm = TRUE),
    economic = mean(economic, na.rm = TRUE),
    colonial= sum(ht_colonial, na.rm=TRUE) / number,
    #asymmetry index
    polity_sd = sd(p_polity2, na.rm = TRUE),
    polyarchy_sd = sd(vdem_polyarchy, na.rm = TRUE),
    libdem_sd=sd(vdem_libdem, na.rm=TRUE),
    partipdem_sd=sd(vdem_partipdem, na.rm=TRUE),
    econ_sd = sd(wdi_log_gdpcapcon2015, na.rm = TRUE),
    # HH index
    hh_poly = sum(poly_share^2, na.rm = TRUE), 
    hh_gdp = sum(gdp_share^2, na.rm = TRUE)
  )

```

---

# merging with MIA data
```{r}
MIA <- MIA %>%
  dplyr::select(ionumber, year, inception, typeI, pooling, delegation, delconstit, poolconstit, DSM, DS_sum_st)%>%
  dplyr::rename(cow_igocode = ionumber)

igo_dataset <- MIA %>% dplyr::left_join(igo, by = c("cow_igocode", "year"))
```

---

# Data Analysis

## Trend of International Organizations Over Time
```{r, results='asis'}

igo_analysis <- igo_dataset%>%
  #dplyr::mutate(AIGO = ifelse(polyarchy < 0.5, 1, 0),
   #             DIGO = ifelse(polyarchy >=0.5, 1, 0)) %>%
  dplyr::mutate_all(~ifelse(is.nan(.), NA, .)) %>%
  dplyr::arrange(ioname, year) %>%
  dplyr::select(ioname, year, polyarchy, everything())%>%
  as.data.frame()

datasummary(All(igo_analysis) ~ Mean + SD + Min + Max, data = igo_analysis) %>%
  kable(caption = "Data Summary") %>%
  kable_styling(font_size = 12, latex_options = "scale_down", full_width = FALSE)

```

---

## Automate the process
```{r}
dem_measure <- function(data, measure_var, threshold = 0.5) {
  processed_data <- data %>%
    mutate(AIGO = ifelse({{measure_var}} < threshold, 1, 0),
           DIGO = ifelse({{measure_var}} >= threshold, 1, 0)) %>%
    mutate_all(~ ifelse(is.nan(.), NA, .)) %>%
    arrange(ioname, year) %>%
    select(ioname, year, {{measure_var}}, everything()) %>%
    as.data.frame()%>%
    group_by(year) %>%
    summarise(DIGO = sum(DIGO, na.rm=TRUE),
            AIGO = sum(AIGO, na.rm=TRUE))
    }
  

#analysis_polyarchy <- dem_measure(igo_dataset, polyarchy, threshold = 0.5)
#analysis_percentage <- dem_measure(igo_dataset, percentage, threshold = 0.5)
#analysis_polity <- dem_measure(igo_dataset, polity, threshold = 0)
#analysis_libdem <- dem_measure(igo_dataset, libdem, threshold = 0.5)
```
