---
title: "DSM and Heterogeneity modeling"
author: "Jihyeon bae"
date: "2024-01-29"
output: html_document
---


# Introduction

## Background and Goals

- Constructing indicators to identify heterogeneity among an IGO's member states based on their regime types. 
- Estimating the correlation between an IGO's political heterogeneity among members and its institutional design choice of Dispute Settlement Mechanism (DSM).


# Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("pacman")
p_load(plyr, dplyr, tidyr, ggplot2, tidyverse, RColorBrewer, readxl,
       readr, haven, countrycode,
       rqog, igoR, modelsummary, knitr, kableExtra, flextable, plm)

```

# Loading processed data
```{r, echo=FALSE, warning=FALSE}
igo_analysis<-readRDS("~/Desktop/Regime-Heterogeneity-and-DSM/data/processed/igo_analysis.RDS")
igo_analysis<-igo_analysis%>%filter(year>=1970 & year <=2015)
```


# Data Analysis

## Trend of International Organizations Over Time
```{r, results='html', eval=FALSE}

datasummary(All(igo_analysis) ~ Mean + SD + Min + Max, data = igo_analysis) %>%
  kable(caption = "Data Summary") %>%
  kable_styling(font_size = 12, latex_options = "scale_down", full_width = FALSE)

```

## Descriptive Visualization
```{r warning=FALSE}

library(gplots)

# Create a plot for each year
plot_year <- plotmeans(DSM ~ year, main = "Heterogeneity across IGOs", data = igo_analysis)

# Save the plot as a PNG file with adjustable size
png("plot_year.png", width = 10, height = 6, units = "in", res = 300)
print(plot_year)
dev.off()

# Create a plot for each IGO
plot_ioname <- plotmeans(DSM ~ ioname, main = "Average DSM score for IGOs", data = igo_analysis,
                         n.label=FALSE, text.n.label="", las = 2)

# Save the plot as a PNG file with adjustable size
png("plot_ioname.png", width = 24, height = 8, units = "in", res = 300)
print(plot_ioname)
dev.off()


```


## Automate the process for other measures of democracy
```{r}
dem_measure <- function(data, measure_var, threshold = 0.5) {
  processed_data <- data %>%
    mutate(AIGO = ifelse({{measure_var}} < threshold, 1, 0),
           DIGO = ifelse({{measure_var}} >= threshold, 1, 0)) %>%
    mutate_all(~ ifelse(is.nan(.), NA, .)) %>%
    arrange(ioname, year) %>%
    select(ioname, year, {{measure_var}}, everything()) %>%
    as.data.frame()%>%
    group_by(year) %>%
    summarise(DIGO = sum(DIGO, na.rm=TRUE),
            AIGO = sum(AIGO, na.rm=TRUE))
    }
  

#analysis_polyarchy <- dem_measure(igo_dataset, polyarchy, threshold = 0.5)
#analysis_percentage <- dem_measure(igo_dataset, percentage, threshold = 0.5)
#analysis_polity <- dem_measure(igo_dataset, polity, threshold = 0)
#analysis_libdem <- dem_measure(igo_dataset, libdem, threshold = 0.5)
```

# Modeling Part 1 - Baseline for model 0, 1 (no hh), model 2 is full extension with hh variables. 
```{r}
library(lmtest)
library(plm)

model0 <-plm(DSM
            ~ polyarchy + polyarchy_sd + econ_sd,
            data = igo_analysis,
            index = c("year"),
            model = "within",
            effect = "time"
)
#summary(model0, vcovBK(model0))

model1 <-plm(DSM
            ~ polyarchy + polyarchy_sd + econ_sd + 
              pooling + 
              ciri_injud + assn + vdem_jucorrdc +
              gdp_cap + population + trade +
               globalization_political + 
              alliances + ptsa + 
              colonial + number + political + social + 1,
            data = igo_analysis,
              index = c("year"),
            model = "within",
            effect = "time"
)
#summary(model1, vcovBK(model1))


#serial correlation
library(tseries)
filtered_data<-igo_analysis%>%filter(!is.na(DSM))
panel.set<-plm.data(filtered_data, index=c("cow_igocode", "year"))
adf.test(panel.set$DSM)
#no unit roots are present here. 

bptest(DSM~polyarchy_sd + factor(cow_igocode), data=panel.set, studentize=F)
#Heteroskedasticity detected. 

##Model 2 has HH variables
model2 <-plm(DSM
            ~ polyarchy + polyarchy_sd + econ_sd + 
              hh_poly + hh_gdp +
              pooling + 
              ciri_injud + assn + vdem_jucorrdc +
              gdp_cap + population + trade +
               globalization_political + 
              alliances + ptsa + 
              colonial + number + political + social + 1,
            data = igo_analysis,
            index = c("year"),
            model = "within",
            effect = "time"
)
summary(model2, vcovBK(model2))

```


## Visualizing models 
```{r}
models1 <- list("Baseline" = model0, "Model 1"=model1, "Model 2"= model2) 

result <- modelsummary(
  models1,
  output = 'flextable',
  stars = TRUE,
  vcov = list(vcovBK(model0), vcovBK(model1), vcovBK(model2)),
  coef_omit = 'Interc',
  inline_css = TRUE,
  bold_italicize_data = TRUE,
  italicize = TRUE,
  font_size = "12pt",  
  override_css = list("body" = "font-family: 'Times New Roman';")
)

result %>%  # text and background color
    color(3, color = 'red') %>%
    #bg(5, bg = 'lightblue') %>%
    # column widths
    autofit()

cm <- c('polyarchy' = 'Polyarchy',
        'polyarchy_sd' = 'Political Heterogeneity', 
        'econ_sd' = 'Economic Heterogeneity', 
        'pooling' = 'Pooling',
        'ciri_injud' = 'Judiciary', 
        'assn' = 'Association', 
        'vdem_jucorrdc' = 'Judicial Corruption',
        'gdp_cap' = 'GDP per capita',
        'population' = 'Population',
        'trade' = 'Trade',
        'globalization_political' = 'Political Globalization',
        'alliances' = 'Alliances',
        'ptsa' = 'Political Terror',
        'colonial' = 'Colonial',
        'number' = 'Number',
        'political' = 'Political',
        'social' = 'Social')


modelplot(models1, coef_omit = 'Interc', coef_map = cm) +
    theme(text = element_text(family = "Times New Roman", size = 14))

# Dropping HH 
modelplot(models1, coef_omit = c(18,19)) +
    theme(text = element_text(family = "Times New Roman", size = 14))

```






# Robustness check
```{r, eval = TRUE}
libdem <-plm(DSM
            ~ libdem + libdem_sd + econ_sd + 
              hh_poly + hh_gdp +
              pooling + 
              ciri_injud + assn + vdem_jucorrdc +
              gdp_cap + population + trade +
               globalization_political + 
              alliances + ptsa + 
              colonial + number + political + social + 1,
            data = igo_analysis,
            index = c("cow_igocode"),
            model = "within",
            effect = "time"
)

summary(libdem, vcovBK(libdem))

particip <-plm(DSM
            ~ partipdem + partipdem_sd +econ_sd + 
             hh_poly + hh_gdp +
              pooling + 
              ciri_injud + assn + vdem_jucorrdc +
              gdp_cap + population + trade +
               globalization_political + 
              alliances + ptsa + 
              colonial + number + political + social + 1,
            data = igo_analysis,
            index = c("year"),
            model = "within",
            effect = "time"
)



robust_models<- list("Liberal"=libdem, "Participatory"=particip) 


robust_result <- modelsummary(
  robust_models,
  output = 'flextable',
  stars = TRUE,
  vcov = list(vcovBK(libdem), vcovBK(particip)),
  coef_omit = 'Interc',
  inline_css = TRUE,
  bold_italicize_data = TRUE,
  italicize = TRUE,
  font_size = "12pt",  
  override_css = list("body" = "font-family: 'Times New Roman';")
)

robust_result %>%  # text and background color
    color(c(3, 41), color = 'red') %>%
    # bg(5, bg = 'lightblue') %>%
    # column widths
    autofit()

```

# Comparing Models and Missingness
```{r, eval=FALSE}
library(naniar)

gg_miss_upset(igo_analysis)

ggplot(igo_analysis, aes(x=cspf_sfi, y=DSM))+geom_miss_point()

ggplot(igo_analysis, aes(x=cspf_sfi, y=DSM))+geom_miss_point()+facet_wrap(~typeI)

gg_miss_fct(igo_analysis, year)
```
