---
title: "DSM and Heterogeneity modeling"
author: "Jihyeon bae"
date: "2024-01-03"
output: html_document
---


# Introduction

## Background and Goals

- Authoritarian Intergovernmental Organizations (AIGO): formal IGOs that largely comprise authoritarian states 
  - Creating aggregate level variable based on country-level data
  - Creating indices for robustness check

- Finding distinctive patterns of AIO compared to democratic counterparts

# Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("pacman")
p_load(plyr, dplyr, tidyr, ggplot2, tidyverse, RColorBrewer, readxl,
       readr, haven, countrycode,
       rqog, igoR, modelsummary, knitr, kableExtra, flextable)

```

# Loading processed data
```{r, echo=FALSE, warning=FALSE}
igo_analysis<-readRDS("igo_analysis.RDS")
```

---

# Data Analysis

## Trend of International Organizations Over Time
```{r, results='html', eval=FALSE}

datasummary(All(igo_analysis) ~ Mean + SD + Min + Max, data = igo_analysis) %>%
  kable(caption = "Data Summary") %>%
  kable_styling(font_size = 12, latex_options = "scale_down", full_width = FALSE)

```

---

## Automate the process
```{r}
dem_measure <- function(data, measure_var, threshold = 0.5) {
  processed_data <- data %>%
    mutate(AIGO = ifelse({{measure_var}} < threshold, 1, 0),
           DIGO = ifelse({{measure_var}} >= threshold, 1, 0)) %>%
    mutate_all(~ ifelse(is.nan(.), NA, .)) %>%
    arrange(ioname, year) %>%
    select(ioname, year, {{measure_var}}, everything()) %>%
    as.data.frame()%>%
    group_by(year) %>%
    summarise(DIGO = sum(DIGO, na.rm=TRUE),
            AIGO = sum(AIGO, na.rm=TRUE))
    }
  

#analysis_polyarchy <- dem_measure(igo_dataset, polyarchy, threshold = 0.5)
#analysis_percentage <- dem_measure(igo_dataset, percentage, threshold = 0.5)
#analysis_polity <- dem_measure(igo_dataset, polity, threshold = 0)
#analysis_libdem <- dem_measure(igo_dataset, libdem, threshold = 0.5)
```


---

## Difference in Pattern?
```{r}
library(lmtest)
library(plm)

pooled<-lm(DSM~ polyarchy + polyarchy_sd +
              econ_sd + gdp_cap + alliances + number + trade + political + social,
            data = igo_analysis)

model0 <-plm(DSM
            ~ polyarchy + polyarchy_sd +
              econ_sd,
            data = igo_analysis,
            index = c("year"),
            model = "within",
            effect = "time"
)
#summary(model0, vcovBK(model0))

model1 <-plm(DSM
            ~ polyarchy + polyarchy_sd +
              hh_poly +
              hh_gdp + econ_sd + pooling + 
              gdp_cap + population + trade +
              globalization + globalization_political + 
              fh_rol + wbgi_pve + 
              cspf_sfi + alliances + ptsa + 
              colonial + 
              number + political + social + 1,
            data = igo_analysis,
            index = c("year", "cow_igocode"),
            model = "within",
            effect = "twoways"
)
#summary(model1, vcovBK(model1))


#serial correlation
library(tseries)
filtered_data<-igo_analysis%>%filter(!is.na(DS_sum_st))
panel.set<-plm.data(filtered_data, index=c("cow_igocode", "year"))
adf.test(panel.set$DS_sum_st)
#no unit roots are present here. 

bptest(DSM~polyarchy_sd + factor(cow_igocode), data=panel.set, studentize=F)
#Heteroskedasticity detected. 

model2 <-plm(DSM
            ~ polyarchy + polyarchy_sd +
              hh_poly +
              hh_gdp + econ_sd + pooling + 
              gdp_cap + population + trade +
              globalization + globalization_political + 
              fh_rol + wbgi_pve + 
              cspf_sfi + alliances + ptsa + 
              colonial + 
              number + political + social + 1,
            data = igo_analysis,
            index = c("year"),
            model = "within",
            effect = "time"
)
#summary(model2, vcovBK(model2))
```


## Visualization of data itself
```{r warning=FALSE}
library(gplots)
plotmeans(DS_sum_st~year, main="Heterogeneity across IGOs", data=igo_analysis)

plotmeans(DS_sum_st~ioname, main="Heterogeneity across IGOs", data=igo_analysis)



```

## Visualizing models
```{r}

models <- list("Two-way"=model1, "Time FE"=model2) 

result <- modelsummary(
  models,
  output = 'flextable',
  stars = TRUE,
  vcov = list(vcovBK(model1), vcovBK(model2)),
  coef_omit = 'Interc',
  inline_css = TRUE,
  bold_italicize_data = TRUE,
  italicize = TRUE,
  font_size = "12pt",  
  override_css = list("body" = "font-family: 'Times New Roman';")
)

result %>%  # text and background color
    color(3, color = 'red') %>%
    #bg(5, bg = 'lightblue') %>%
    # column widths
    autofit()


```

# Robustness check
```{r}
percentage <-plm(DSM
            ~ percentage + polyarchy_sd +
              hh_poly +
              hh_gdp + econ_sd + pooling + 
              gdp_cap + population + trade +
              globalization + globalization_political + 
              fh_rol + wbgi_pve + 
              cspf_sfi + alliances + ptsa + 
              colonial + 
              number + political + social + 1,
            data = igo_analysis,
            index = c("year", "cow_igocode"),
            model = "within",
            effect = "twoway"
)
# summary(percentage, vcovBK(percentage))

percentage2 <-plm(DSM
            ~ percentage + polyarchy_sd +
              hh_poly +
              hh_gdp + econ_sd + pooling + 
              gdp_cap + population + trade +
              globalization + globalization_political + 
              fh_rol + wbgi_pve + 
              cspf_sfi + alliances + ptsa + 
              colonial + 
              number + political + social + 1,
            data = igo_analysis,
            index = c("year"),
            model = "within",
            effect = "time"
)

#summary(percentage2, vcovBK(percentage2))


robust_models<- list("Two-way"=percentage, "Time FE"=percentage2) 


robust_result <- modelsummary(
  robust_models,
  output = 'flextable',
  stars = TRUE,
  vcov = list(vcovBK(percentage), vcovBK(percentage2)),
  coef_omit = 'Interc',
  inline_css = TRUE,
  bold_italicize_data = TRUE,
  italicize = TRUE,
  font_size = "12pt",  
  override_css = list("body" = "font-family: 'Times New Roman';")
)

robust_result %>%  # text and background color
    color(3, color = 'red') %>%
    # bg(5, bg = 'lightblue') %>%
    # column widths
    autofit()
```
