---
title: "DSM and Heterogeneity modeling"
author: "Jihyeon bae"
date: "2024-01-29"
output: html_document
---


# Introduction

## Background and Goals

- Constructing indicators to identify heterogeneity among an IGO's member states based on their regime types. 
- Estimating the correlation between an IGO's political heterogeneity among members and its institutional design choice of Dispute Settlement Mechanism (DSM).


# Setup and packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("pacman")
p_load(plyr, dplyr, tidyr, ggplot2, tidyverse, RColorBrewer, readxl,
       readr, haven, countrycode,
       rqog, igoR, modelsummary, knitr, kableExtra, flextable, plm)

```

# Loading Raw data
```{r, echo=FALSE, warning=FALSE}

#Qualities of Government
qog<-rqog::read_qog(which_data = "standard", data_type = "time-series") 

qog<-qog%>%
  filter(year>=1950 & year<2015) 

#Measuring International Authority
MIA <- read.csv("~/Desktop/Regime-Heterogeneity-and-DSM/data/raw/MIA_updated.csv")

#Correlates of War
COW<-igoR::igo_year_format3 %>%
  filter(year>=1950)
```

---

# Country-level variables

```{r, echo = FALSE}
country <-qog %>%
  dplyr::select(cname, year, ccodecow, 
         # World Development Indicators
          wdi_gdpcapcon2015, wdi_pop, wdi_trade, 
         # judicial corruption decision
          vdem_jucorrdc, 
          ciri_assn, #freedom of association
          ciri_injud, # independence of the judiciary
        # Security variables
          atop_number, 
        # Democracy Indices 
          vdem_polyarchy, vdem_libdem, vdem_partipdem,
        # Democracy Indices (Alternatives for robustness check)
          p_polity2, bmr_dem, ht_regtype1,
        # Globalization Index (political)
           kofgi_dr_pg=dr_pg,
        # colonial legacy
           ht_colonial,
        # Political Terror Scale
         pts_ptsa=gd_ptsa) %>%  
  dplyr::mutate(wdi_log_gdpcapcon2015=log(wdi_gdpcapcon2015),
         wdi_log_pop=log(wdi_pop),
         wdi_log_trade=log(wdi_trade))%>%
  dplyr::rename(ccode=ccodecow)%>%
  dplyr::relocate(ccode, cname, year)

country <- country %>%
  mutate(ht_colonial=ifelse(ht_colonial>1, 1, 0),
         ht_regtype1=case_when(ht_regtype1==9 ~ 5, 
                            ht_regtype1==99 ~ 6,
                            ht_regtype1==100 ~ 7,
                            TRUE ~ ht_regtype1))

```

```{r, echo = FALSE}
#Check rows that have NA values for VARIABLE_OF_INTEREST
country[is.na(country$ccode),]

#Fill in NAs
country$ccode[country$cname == "Ethiopia"] <- 530
country$ccode[country$cname == "Germany"] <- 255
country$ccode[country$cname == "Yemen Democratic"] <- 680
country$ccode[country$cname == "Yemen"] <- 679
country$ccode[country$cname == "Sudan"] <- 625
country$ccode[country$cname == "Cyprus"] <- 352
country$ccode[country$cname == "Pakistan"] <- 770
country$ccode[country$cname == "Vietnam, North"] <- 816
country$ccode[country$cname == "Vietnam, South"] <- 817
country$cname[country$cname == "USSR"] <- "Russian Federation (the)"


#Run below code to check year-availability for VARIABLE_OF_INTEREST
#summary(country[!is.na(country$VARIABLE_OF_INTEREST),])

country<-country%>%
  filter(!is.na(ccode))
```

---

## Transposing IGO-level data into country-level data

- Pivot longer so that row represents country
```{r}
COW <- COW %>%
  dplyr::rename(cow_igocode = ionum)%>%
  dplyr::select(-c(igocode, version, accuracyofpre1965membershipdates,sourcesandnotes, imputed)) %>%
  dplyr::relocate(cow_igocode, ioname, year, political, social, economic)%>%
  pivot_longer(c(`afghanistan`:`zimbabwe`),
                      names_to="country",
                      values_to="membership")%>%
  dplyr::filter(membership==1) #member states only

```

## Identify ambiguous country names
```{r}
COW <- COW %>%
  mutate(country = recode(country, "austriahungary" = "Austria-Hungary",
                          "domrepublic"="Dominican Republic",
                          "etimor"="East Timor",
                          "hessegrand"="Hesse Grand Ducal",
                          "micronesiafs"="Federated States of Micronesia",
                          "nokorea"="North Korea",
                          "soafrica"="South Africa",
                          "sokorea"="South Korea",
                          "stlucia"="St. Lucia",
                          "wgermany"="German Federal Republic",
                          "syemen"="Yemen People's Republic",
                          .default = country))

#Attaching Country numeric code to character values
COW$ccode<-countrycode(COW$country, 
              origin='country.name', 
              destination='cown', 
              warn = TRUE)

COW<-COW%>%
  filter(!is.na(ccode))
```


## Extracting member-state information for each IGO
```{r}
igo_master <- COW %>%
  dplyr::inner_join(country, by=c("ccode", "year"))%>%
  dplyr::select(-c(orgname, longorgname, membership))


igo_master <- igo_master %>%
  group_by(cow_igocode, year)%>%
  mutate(gdp_share= wdi_gdpcapcon2015/sum(wdi_gdpcapcon2015, na.rm=TRUE),
         poly_share=vdem_polyarchy/sum(vdem_polyarchy, na.rm=TRUE))

```



# Variable Generation
## summarising state level info to igo level
```{r}
igo <- igo_master %>%
  dplyr::group_by(cow_igocode, ioname, year) %>%
  dplyr::summarise(
    #average democracy scores
    polyarchy = mean(vdem_polyarchy, na.rm = TRUE),
    w.polyarchy = weighted.mean(vdem_polyarchy, w = wdi_log_gdpcapcon2015/sum(wdi_log_gdpcapcon2015), na.rm=TRUE),
    polyarchy_median = median(vdem_polyarchy, na.rm = TRUE),
    partipdem = mean(vdem_partipdem, na.rm=TRUE),
    partipdem_median = median(vdem_partipdem, na.rm=TRUE),
    libdem = mean(vdem_libdem, na.rm = TRUE),
    libdem_median = median(vdem_libdem, na.rm = TRUE),
    #economic variables
    gdp_cap = mean(wdi_log_gdpcapcon2015, na.rm = TRUE),
    population=mean(wdi_log_pop, na.rm=TRUE),
    trade=mean(wdi_log_trade, na.rm=TRUE),
    globalization=mean(kofgi_dr_pg, na.rm=TRUE),
    alliances = mean(atop_number, na.rm = TRUE),
    # Governance
    vdem_jucorrdc=mean(vdem_jucorrdc, na.rm = TRUE),
    ciri_injud=mean(ciri_injud, na.rm=TRUE),
    assn=mean(ciri_assn, na.rm=TRUE),
    ptsa=mean(pts_ptsa, na.rm=TRUE), 
    #characteristics 
    number = n(),
    trade = mean(wdi_log_trade, na.rm = TRUE),
    percentage = sum(bmr_dem, na.rm = TRUE) / number,
    political = mean(political, na.rm = TRUE),
    social = mean(social, na.rm = TRUE),
    economic = mean(economic, na.rm = TRUE),
    colonial= sum(ht_colonial, na.rm=TRUE) / number,
    #asymmetry index
    polity_sd = sd(p_polity2, na.rm = TRUE),
    polyarchy_sd = sd(vdem_polyarchy, na.rm = TRUE),
    libdem_sd=sd(vdem_libdem, na.rm=TRUE),
    partipdem_sd=sd(vdem_partipdem, na.rm=TRUE),
    econ_sd = sd(wdi_log_gdpcapcon2015, na.rm = TRUE),
    # HH index
    hh_poly = sum(poly_share^2, na.rm = TRUE)- (1/number), 
    hh_gdp = sum(gdp_share^2, na.rm = TRUE)- (1/number)
)

  
```

---

# merging with MIA data
```{r}
MIA <- MIA %>%
  dplyr::select(ionumber, year, inception, typeI, pooling, delegation, delconstit, poolconstit, DSM)%>%
  dplyr::rename(cow_igocode = ionumber)

igo_dataset <- MIA %>% dplyr::left_join(igo, by = c("cow_igocode", "year"))
saveRDS(igo_dataset, "~/Desktop/Regime-Heterogeneity-and-DSM/data/processed/data_May2024.RDS")
```


# Data Analysis

## Trend of International Organizations Over Time
```{r, results='html', eval=FALSE}

datasummary(All(igo_dataset) ~ Mean + SD + Min + Max, data = igo_dataset) %>%
  kable(caption = "Data Summary") %>%
  kable_styling(font_size = 12, latex_options = "scale_down", full_width = FALSE)

```

## Descriptive Visualization
```{r warning=FALSE}

library(gplots)

# Create a plot for each year
plot_year <- plotmeans(DSM ~ year, main = "Heterogeneity across IGOs", data = igo_dataset)

# Save the plot as a PNG file with adjustable size
png("plot_year.png", width = 10, height = 6, units = "in", res = 300)
print(plot_year)
dev.off()

# Create a plot for each IGO
plot_ioname <- plotmeans(DSM ~ ioname, main = "Average DSM score for IGOs", data = igo_dataset,
                         n.label=FALSE, text.n.label="", las = 2)

# Save the plot as a PNG file with adjustable size
png("plot_ioname.png", width = 24, height = 8, units = "in", res = 300)
print(plot_ioname)
dev.off()


```

# Modeling Part 1 - Baseline for model 0, 1 (no hh), model 2 is full extension with hh variables. 
```{r}
library(lmtest)
library(plm)

model0 <-plm(DSM
            ~ polyarchy + polyarchy_sd + econ_sd,
            data = igo_dataset,
            index = c("year"),
            model = "within",
            effect = "time"
)
#summary(model0, vcovBK(model0))

model1 <-plm(DSM
            ~ polyarchy + polyarchy_sd + econ_sd + 
              pooling + 
              ciri_injud + assn + vdem_jucorrdc +
              gdp_cap + population + trade +
               globalization + 
              alliances + ptsa + 
              colonial + number + political + social + 1,
            data = igo_dataset,
              index = c("year"),
            model = "within",
            effect = "time"
)
summary(model1, vcovBK(model1))


#serial correlation
library(tseries)
filtered_data<-igo_dataset%>%filter(!is.na(DSM))
panel.set<-plm.data(filtered_data, index=c("cow_igocode", "year"))
adf.test(panel.set$DSM)
#no unit roots are present here. 

bptest(DSM~polyarchy_sd + factor(cow_igocode), data=panel.set, studentize=F)
#Heteroskedasticity detected. 

##Model 2 has HH variables
model2 <-plm(DSM
            ~ polyarchy + polyarchy_sd + econ_sd + 
              hh_poly + hh_gdp +
              pooling + 
              ciri_injud + assn + vdem_jucorrdc +
              gdp_cap + population + trade +
               globalization + 
              alliances + ptsa + 
              colonial + number + political + social + 1,
            data = igo_dataset,
            index = c("year"),
            model = "within",
            effect = "time"
)
summary(model2, vcovBK(model2))

```


## Visualizing models 
```{r}
models1 <- list("Baseline" = model0, "Model 1"=model1, "Model 2"= model2) 

result <- modelsummary(
  models1,
  output = 'flextable',
  stars = TRUE,
  vcov = list(vcovBK(model0), vcovBK(model1), vcovBK(model2)),
  coef_omit = 'Interc',
  inline_css = TRUE,
  bold_italicize_data = TRUE,
  italicize = TRUE,
  font_size = "12pt",  
  override_css = list("body" = "font-family: 'Times New Roman';")
)

result %>%  # text and background color
    color(3, color = 'red') %>%
    #bg(5, bg = 'lightblue') %>%
    # column widths
    autofit()

cm <- c('hh_poly' = 'Political HHI',
        'hh_gdp' = 'Economic HHI',
        'social' = 'Social',
        'political' = 'Political',
        'colonial' = 'Colonial',
        'ptsa' = 'Political Terror',
        'alliances' = 'Alliances',
        'globalization' = 'Political Globalization',
        'trade' = 'Trade',
        'population' = 'Population',
        'gdp_cap' = 'GDP per capita',
        'vdem_jucorrdc' = 'Judicial Corruption',
        'assn' = 'Association', 
        'ciri_injud' = 'Judiciary', 
        'pooling' = 'Pooling',
        'econ_sd' = 'Economic Heterogeneity', 
        'polyarchy_sd' = 'Political Heterogeneity', 
        'polyarchy' = 'Polyarchy')

models <- list("Model 1"=model1, "Model 2"= model2) 


modelplot(models, coef_omit = 'Interc', coef_map = cm) +
    theme(text = element_text(family = "Times New Roman", size = 14))+
    scale_x_continuous(limits=c(-1,1))+
    aes(shape=model)+
    scale_shape_manual(values=c(17, 16))+
    scale_color_grey(start=0.2, end=0.6)

# Dropping HH 

cm2 <- c('social' = 'Social',
        'political' = 'Political',
        'colonial' = 'Colonial',
        'ptsa' = 'Political Terror',
        'alliances' = 'Alliances',
        'globalization' = 'Political Globalization',
        'trade' = 'Trade',
        'population' = 'Population',
        'gdp_cap' = 'GDP per capita',
        'vdem_jucorrdc' = 'Judicial Corruption',
        'assn' = 'Association', 
        'ciri_injud' = 'Judiciary', 
        'pooling' = 'Pooling',
        'econ_sd' = 'Economic Heterogeneity', 
        'polyarchy_sd' = 'Political Heterogeneity', 
        'polyarchy' = 'Polyarchy')

modelplot(models, coef_omit = 'Interc', coef_map = cm2)+
    theme(text = element_text(family = "Times New Roman", size = 14))+
    scale_x_continuous(limits=c(-1,1))+
    aes(shape=model)+
    scale_shape_manual(values=c(17, 16))+
    scale_color_grey(start=0.2, end=0.6)

```






# Robustness check
```{r, eval = TRUE}
libdem <-plm(DSM
            ~ libdem + libdem_sd + econ_sd + 
              hh_poly + hh_gdp +
              pooling + 
              ciri_injud + assn + vdem_jucorrdc +
              gdp_cap + population + trade +
               globalization + 
              alliances + ptsa + 
              colonial + number + political + social + 1,
            data = igo_dataset,
            index = c("cow_igocode"),
            model = "within",
            effect = "time"
)

summary(libdem, vcovBK(libdem))

particip <-plm(DSM
            ~ partipdem + partipdem_sd +econ_sd + 
             hh_poly + hh_gdp +
              pooling + 
              ciri_injud + assn + vdem_jucorrdc +
              gdp_cap + population + trade +
               globalization + 
              alliances + ptsa + 
              colonial + number + political + social + 1,
            data = igo_dataset,
            index = c("year"),
            model = "within",
            effect = "time"
)



robust_models<- list("Liberal"=libdem, "Participatory"=particip) 


robust_result <- modelsummary(
  robust_models,
  output = 'flextable',
  stars = TRUE,
  vcov = list(vcovBK(libdem), vcovBK(particip)),
  coef_omit = 'Interc',
  inline_css = TRUE,
  bold_italicize_data = TRUE,
  italicize = TRUE,
  font_size = "12pt",  
  override_css = list("body" = "font-family: 'Times New Roman';")
)

robust_result %>%  # text and background color
    color(c(3, 41), color = 'red') %>%
    # bg(5, bg = 'lightblue') %>%
    # column widths
    autofit()

```

# Comparing Models and Missingness
```{r, eval=FALSE}
library(naniar)

gg_miss_upset(igo_dataset)

ggplot(igo_dataset, aes(x=cspf_sfi, y=DSM))+geom_miss_point()

ggplot(igo_dataset, aes(x=cspf_sfi, y=DSM))+geom_miss_point()+facet_wrap(~typeI)

gg_miss_fct(igo_dataset, year)
```
